---
title: "Assignment 2 FML"
Author: Adithya Reddy Mettu
date: "2024-02-26"
output:
  word_document: default
  pdf_document: default
  html_document: default
---
***

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



***

###Data Import and Cleaning

First, load the required libraries
```{r}
library(ggplot2)
library(lattice)
library(class)
library(caret)

```

## Loading required package: ggplot2
## Loading required package: lattice
```{r}
library(e1071)

```
#Read the data.
```{r}
universal.df <- read.csv("C:/Users/adith/Downloads/UniversalBank.csv")
dim(universal.df)

t(t(names(universal.df))) # The t function creates a transpose of the dataframe

```

#Drop ID and ZIP

```{r}
universal.df <- universal.df[,-c(1,5)]

```

#Split Data into 60% training and 40% validation. There are many ways to do this. We will look at 2 different ways. Before we split, let us transform categorical variables into dummy variables

# Only Education needs to be converted to factor
```{r}
universal.df$Education <- as.factor(universal.df$Education)

```

# Now, convert Education to Dummy Variables
```{r}
groups <- dummyVars(~., data = universal.df) # This creates the dummy groups
universal_m.df <- as.data.frame(predict(groups,universal.df))


set.seed(1)  # Important to ensure that we get the same sample if we rerun the code
train.index <- sample(row.names(universal_m.df), 0.6*dim(universal_m.df)[1])
valid.index <- setdiff(row.names(universal_m.df), train.index)  
train.df <- universal_m.df[train.index,]
valid.df <- universal_m.df[valid.index,]
t(t(names(train.df)))

```

```{r}

library(caTools)
set.seed(1)
split <- sample.split(universal_m.df, SplitRatio = 0.6)
training_set <- subset(universal_m.df, split == TRUE)
validation_set <- subset(universal_m.df, split == FALSE)

# Print the sizes of the training and validation sets
print(paste("The size of the training set is:", nrow(training_set)))
## [1] "The size of the training set is: 2858"
print(paste("The size of the validation set is:", nrow(validation_set)))
## [1] "The size of the validation set is: 2142"
#Now, let us normalize the data

train.norm.df <- train.df[,-10] # Note that Personal Income is the 10th variable
valid.norm.df <- valid.df[,-10]

norm.values <- preProcess(train.df[, -10], method=c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])

```
***
###Questions

Consider the following customer:

Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1, and Credit Card = 1. Perform a k-NN classification with all predictors except ID and ZIP code using k = 1. Remember to transform categorical predictors with more than two categories into dummy variables first. Specify the success class as 1 (loan acceptance), and use the default cutoff value of 0.5. How would this customer be classified?
# We have converted all categorical variables to dummy variables
# Let's create a new sample

```{r}
new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education.1 = 0,
  Education.2 = 1,
  Education.3 = 0,
  Mortgage = 0,
  Securities.Account = 0,
  CD.Account = 0,
  Online = 1,
  CreditCard = 1
)

```

# Normalize the new customer
```{r}
new.cust.norm <- new_customer
new.cust.norm <- predict(norm.values, new.cust.norm)
```

#Now,let us predict using knn

```{r}
knn.pred1 <- class::knn(train = train.norm.df, 
                       test = new.cust.norm, 
                       cl = train.df$Personal.Loan, k = 1)
knn.pred1
```

What is a choice of k that balances between overfitting and ignoring the predictor information?
# Calculate the accuracy for each value of k
# Set the range of k values to consider
```{r}
accuracy.df <- data.frame(k = seq(1, 15, 1), overallaccuracy = rep(0, 15))
for(i in 1:15) {
  knn.pred <- class::knn(train = train.norm.df, 
                         test = valid.norm.df, 
                         cl = train.df$Personal.Loan, k = i)
  accuracy.df[i, 2] <- confusionMatrix(knn.pred, 
                                       as.factor(valid.df$Personal.Loan),positive = "1")$overall[1]
}

which(accuracy.df[,2] == max(accuracy.df[,2]))

```

```{r}
plot(accuracy.df$k,accuracy.df$overallaccuracy)

```
 

 ***

Show the confusion matrix for the validation data that results from using the best k.
# Best k based on validation accuracy
```{r}

best_k <- which(accuracy.df$overallaccuracy == max(accuracy.df$overallaccuracy))

```


# Classify the validation data using the best k
```{r}

knn.pred_best <- class::knn(train = train.norm.df, 
                            test = valid.norm.df, 
                            cl = train.df$Personal.Loan, k = best_k)

```


# Create the confusion matrix
```{r}

conf_matrix <- confusionMatrix(knn.pred_best, 
                               as.factor(valid.df$Personal.Loan), positive = "1")
```

# Display the confusion matrix
```{r}
conf_matrix

```
## Confusion Matrix and Statistics

Consider the following customer: Age = 40, Experience = 10, Income = 84, Family = 2, CCAvg = 2, Education_1 = 0, Education_2 = 1, Education_3 = 0, Mortgage = 0, Securities Account = 0, CD Account = 0, Online = 1 and Credit Card = 1. Classify the customer using the best k.
# Create a data frame for the new customer with the same column names
# Create a data frame for the new customer with matching column names

```{r}
new_customer <- data.frame(
  Age = 40,
  Experience = 10,
  Income = 84,
  Family = 2,
  CCAvg = 2,
  Education_1 = 0,
  Education_2 = 1,
  Education_3 = 0,
  Mortgage = 0,
  `Securities Account` = 0,  # Use backticks for column names with spaces
  `CD Account` = 0,  # Use backticks for column names with spaces
  Online = 1,
  `Credit Card` = 1  # Use backticks for column names with spaces
)

```

# Normalize the new customer data using the same preprocessing
```{r}
new.cust.norm <- predict(norm.values, new.cust.norm)

```

# Predict whether the new customer accepts a loan using the best k
```{r}

new_customer_classification <- class::knn(train = train.norm.df, 
                                          test = new.cust.norm, 
                                          cl = train.df$Personal.Loan, k = best_k)

```

# Display the classification result
```{r}

new_customer_classification

```

Repartition the data, this time into training, validation, and test sets (50% : 30% : 20%). Apply the k-NN method with the k chosen above. Compare the confusion matrix of the test set with that of the training and validation sets. Comment on the differences and their reason.
#Comment on the differences and their reason: Differences in confusion matrices between training, validation, and test sets are expected due to the different roles and characteristics of each set. Discrepancies could indicate potential issues such as overfitting or data sampling differences. It’s crucial to monitor these differences and make adjustments to ensure the model generalizes well to unseen data.

# Set the seed for reproducibility
```{r}

set.seed(1)

```

# Repartition the data into training (50%), validation (30%), and test (20%) sets
```{r}
train.index <- sample(1:nrow(universal_m.df), 0.5 * nrow(universal_m.df))
valid.test.index <- setdiff(1:nrow(universal_m.df), train.index)
valid.index <- sample(valid.test.index, 0.3 * length(valid.test.index))
test.index <- setdiff(valid.test.index, valid.index)

train.df <- universal_m.df[train.index, ]
valid.df <- universal_m.df[valid.index, ]
test.df <- universal_m.df[test.index, ]

```

# Normalize the data for each set
```{r}
norm.values <- preProcess(train.df[, -10], method = c("center", "scale"))
train.norm.df <- predict(norm.values, train.df[, -10])
valid.norm.df <- predict(norm.values, valid.df[, -10])
test.norm.df <- predict(norm.values, test.df[, -10])
```

# Classify the data using the best k
```{r}
knn.pred_train <- class::knn(train = train.norm.df, 
                             test = train.norm.df, 
                             cl = train.df$Personal.Loan, k = best_k)

knn.pred_valid <- class::knn(train = train.norm.df, 
                             test = valid.norm.df, 
                             cl = train.df$Personal.Loan, k = best_k)

knn.pred_test <- class::knn(train = train.norm.df, 
                            test = test.norm.df, 
                            cl = train.df$Personal.Loan, k = best_k)
```

# Create confusion matrices for each set
```{r}
conf_matrix_train <- confusionMatrix(knn.pred_train, 
                                     as.factor(train.df$Personal.Loan), positive = "1")

conf_matrix_valid <- confusionMatrix(knn.pred_valid, 
                                     as.factor(valid.df$Personal.Loan), positive = "1")

conf_matrix_test <- confusionMatrix(knn.pred_test, 
                                    as.factor(test.df$Personal.Loan), positive = "1")

```

# Display the confusion matrices
```{r}
conf_matrix_train

```

## Confusion Matrix and Statistics
```{r}
conf_matrix_valid

```

```{r}

conf_matrix_test

```
